---
title: "Engine Health - Data Preprocessing"
output:
  html_document:
    df_print: paged
Author: Soham Nimale
---

```{r}
ds = read.csv("C:/Users/nimal/eclipse-workspace/CourseProjects/syCourseProject/DS/dataset/engine_data.csv")
```

column details

```{r}
summary(ds)
```
```{r}
str(ds)
```
### Checking For Null Values
```{r}
# checking for null values
sum(is.na(ds$Engine.rpm))
sum(is.na(ds$Coolant.temp))
sum(is.na(ds$lub.oil.temp))
sum(is.na(ds$Coolant.pressure))
sum(is.na(ds$Fuel.pressure))
sum(is.na(ds$Lub.oil.pressure))
sum(is.na(ds$Engine.Condition))
```
## Remove Outliers
#### Engine RPM
```{r}
# Engine RPM analysis
data = ds$Engine.rpm
min(data)
max(data)
mean(data)
```
```{r}
library(ggplot2)
ggplot(data=ds, mapping=aes(x=Engine.rpm))+geom_histogram(bins=100)
```
```{r}
Q1 = quantile(ds$Engine.rpm, 0.25)
Q3 = quantile(ds$Engine.rpm, 0.75)
print(paste("25th Percentile => ", Q1))
print(paste("75th Percentile => ", Q3))
IQR = Q3 - Q1
print(paste("Inter Quartile Range => ", IQR))
print(paste("Theoritical Min => ", Q1 - 1.7 * IQR))
print(paste("Theoritical Max => ", Q3 + 1.7 * IQR))
```
```{r}
nrow(subset(ds, ds$Engine.rpm < Q1 - 1.5 * IQR))
nrow(subset(ds, ds$Engine.rpm > Q3 + 1.5 * IQR))
```
The number of outliers are small enough for us to drop them
```{r}
# important => do not repeat this step again and again, it will try and remove outliers from the data set who already has its outliars removed.
ds = subset(ds, ds$Engine.rpm > Q1 - 1.5 * IQR & ds$Engine.rpm < Q3 + 1.5 * IQR)
```

```{r}
ggplot(data=ds, mapping=aes(x=Engine.rpm))+geom_histogram(bins=100)
```
#### Engine Lub Oil Pressure
```{r}
data = ds$Lub.oil.pressure
min(data)
max(data)
mean(data)
```

```{r}
ggplot(data=ds, mapping=aes(x=Lub.oil.pressure))+geom_histogram(bins=100)
```

```{r}
mean = mean(ds$Lub.oil.pressure)
std = sd(ds$Lub.oil.pressure)
print(mean)
print(std)
```

```{r}
nrow(subset(ds, (ds$Lub.oil.pressure-mean)/std > 3.0))
nrow(subset(ds, (ds$Lub.oil.pressure-mean)/std < -3.0))
print(paste(nrow(ds), " = ", sum((ds$Lub.oil.pressure-mean)/std < -3.0 |  (ds$Lub.oil.pressure-mean)/std > 3.0), " + ", sum((ds$Lub.oil.pressure-mean)/std >= -3.0 &  (ds$Lub.oil.pressure-mean)/std <= 3.0)))
```
The number of outliers are small enough for us to drop them
```{r}
# important => do not repeat this step again and again, it will try and remove outliers from the data set who already has its outliars removed.
ds = subset(ds, (ds$Lub.oil.pressure-mean)/std >= -3.0 &  (ds$Lub.oil.pressure-mean)/std <= 3.0)
```

```{r}
ggplot(data=ds, mapping=aes(x=Lub.oil.pressure))+geom_histogram(bins=100)
```
#### Fuel.pressure
```{r}
data = ds$Fuel.pressure
min(data)
max(data)
mean(data)
```

```{r}
ggplot(data=ds, mapping=aes(x=Fuel.pressure))+geom_histogram(bins=100)
```

```{r}
mean = mean(ds$Fuel.pressure)
std = sd(ds$Fuel.pressure)
print(mean)
print(std)
```

```{r}
nrow(subset(ds, (ds$Fuel.pressure-mean)/std > 3.0))
nrow(subset(ds, (ds$Fuel.pressure-mean)/std < -3.0))
print(paste(nrow(ds), " = ", sum((ds$Fuel.pressure-mean)/std < -3.0 |  (ds$Fuel.pressure-mean)/std > 3.0), " + ", sum((ds$Fuel.pressure-mean)/std >= -3.0 &  (ds$Fuel.pressure-mean)/std <= 3.0)))
```
Hmm, just to be sure we also check out outliers by IQR method
```{r}
Q1 = quantile(ds$Fuel.pressure, 0.25)
Q3 = quantile(ds$Fuel.pressure, 0.75)
print(paste("25th Percentile => ", Q1))
print(paste("75th Percentile => ", Q3))
IQR = Q3 - Q1
print(paste("Inter Quartile Range => ", IQR))
print(paste("Theoritical Min => ", Q1 - 1.7 * IQR))
print(paste("Theoritical Max => ", Q3 + 1.7 * IQR))
```
```{r}
nrow(subset(ds, ds$Fuel.pressure < Q1 - 1.5 * IQR))
nrow(subset(ds, ds$Fuel.pressure > Q3 + 1.5 * IQR))
```
too much loss of data here as expected, so we go with our original plan, i.e. using z score
```{r}
# important => do not repeat this step again and again, it will try and remove outliers from the data set who already has its outliars removed.
ds = subset(ds, (ds$Fuel.pressure-mean)/std >= -3.0 &  (ds$Fuel.pressure-mean)/std <= 3.0)
```

```{r}
ggplot(data=ds, mapping=aes(x=Fuel.pressure))+geom_histogram(bins=100)
```
#### Coolant.pressure
```{r}
data = ds$Coolant.pressure
min(data)
max(data)
mean(data)
```

```{r}
ggplot(data=ds, mapping=aes(x=Coolant.pressure))+geom_histogram(bins=100)
```

```{r}
mean = mean(ds$Coolant.pressure)
std = sd(ds$Coolant.pressure)
print(mean)
print(std)
```

```{r}
nrow(subset(ds, (ds$Coolant.pressure-mean)/std > 3.0))
nrow(subset(ds, (ds$Coolant.pressure-mean)/std < -3.0))
print(paste(nrow(ds), " = ", sum((ds$Coolant.pressure-mean)/std < -3.0 |  (ds$Coolant.pressure-mean)/std > 3.0), " + ", sum((ds$Coolant.pressure-mean)/std >= -3.0 &  (ds$Coolant.pressure-mean)/std <= 3.0)))
```
again, just to be sure, since graph is not perfectly normal, we check IQR method too for this
```{r}
Q1 = quantile(ds$Coolant.pressure, 0.25)
Q3 = quantile(ds$Coolant.pressure, 0.75)
print(paste("25th Percentile => ", Q1))
print(paste("75th Percentile => ", Q3))
IQR = Q3 - Q1
print(paste("Inter Quartile Range => ", IQR))
print(paste("Theoritical Min => ", Q1 - 1.7 * IQR))
print(paste("Theoritical Max => ", Q3 + 1.7 * IQR))
```
```{r}
nrow(subset(ds, ds$Coolant.pressure < Q1 - 1.5 * IQR))
nrow(subset(ds, ds$Coolant.pressure > Q3 + 1.5 * IQR))
```
although this number is more than z score method, we still use this since it seems fair to drop these rows seeing the skeweness of graph on +ve side
```{r}
# important => do not repeat this step again and again, it will try and remove outliers from the data set who already has its outliars removed.
ds = subset(ds, ds$Coolant.pressure > Q1 - 1.5 * IQR & ds$Coolant.pressure < Q3 + 1.5 * IQR)
```

```{r}
ggplot(data=ds, mapping=aes(x=Coolant.pressure))+geom_histogram(bins=100)
```
#### lub.oil.temp
```{r}
data = ds$lub.oil.temp
min(data)
max(data)
mean(data)
```

```{r}
ggplot(data=ds, mapping=aes(x=lub.oil.temp))+geom_histogram(bins=100)
```
# THIS IS A WIERD LOOKING GRAPH, NEED HELP TO REMOVE OUTLIERS HERE, FOR NOW WE CONTINUE WITHOUT REMOVEING OUTLIERS

#### Coolant.temp
```{r}
data = ds$Coolant.temp
min(data)
max(data)
mean(data)
```

```{r}
ggplot(data=ds, mapping=aes(x=Coolant.temp))+geom_histogram(bins=100)
```
again, a wierd looking graph, but something seems odd here :> we will remove outlier manually if needed
```{r}
mean = mean(ds$Coolant.temp)
std = sd(ds$Coolant.temp)
print(mean)
print(std)
```

```{r}
nrow(subset(ds, (ds$Coolant.temp-mean)/std > 3.0))
nrow(subset(ds, (ds$Coolant.temp-mean)/std < -3.0))
print(paste(nrow(ds), " = ", sum((ds$Coolant.temp-mean)/std < -3.0 |  (ds$Coolant.temp-mean)/std > 3.0), " + ", sum((ds$Coolant.temp-mean)/std >= -3.0 &  (ds$Coolant.temp-mean)/std <= 3.0)))
```
since outliers are obvious, we continue with the removal
```{r}
# important => do not repeat this step again and again, it will try and remove outliers from the data set who already has its outliars removed.
ds = subset(ds, (ds$Coolant.temp-mean)/std >= -3.0 &  (ds$Coolant.temp-mean)/std <= 3.0)
```

```{r}
ggplot(data=ds, mapping=aes(x=Coolant.temp))+geom_histogram(bins=100)
```

# Save the outliers and missing values removed File as a checkpoint
```{r}
write.csv(ds, "C:/Users/nimal/eclipse-workspace/CourseProjects/syCourseProject/DS/dataset/engine_data_removed_missing_and_outliers.csv", row.names=FALSE)
```

# feature scaling : now we bring all the columns scale to 0 to 1 (we use normalization since most graphs are not NORMAL, hence z score would be a poor indicator)

```{r}
# get the scaled one from file
ds = read.csv("C:/Users/nimal/eclipse-workspace/CourseProjects/syCourseProject/DS/dataset/engine_data_removed_missing_and_outliers.csv")
head(ds)
```


```{r}
ds$Engine.rpm = (ds$Engine.rpm - min(ds$Engine.rpm))/(max(ds$Engine.rpm) - min(ds$Engine.rpm))
ds$Lub.oil.pressure = (ds$Lub.oil.pressure - min(ds$Lub.oil.pressure))/(max(ds$Lub.oil.pressure) - min(ds$Lub.oil.pressure))
ds$Fuel.pressure = (ds$Fuel.pressure - min(ds$Fuel.pressure))/(max(ds$Fuel.pressure) - min(ds$Fuel.pressure))
ds$Coolant.pressure = (ds$Coolant.pressure - min(ds$Coolant.pressure))/(max(ds$Coolant.pressure) - min(ds$Coolant.pressure))
ds$lub.oil.temp = (ds$lub.oil.temp - min(ds$lub.oil.temp))/(max(ds$lub.oil.temp) - min(ds$lub.oil.temp))
ds$Coolant.temp = (ds$Coolant.temp - min(ds$Coolant.temp))/(max(ds$Coolant.temp) - min(ds$Coolant.temp))
```

#### test the scaled values
```{r}
print(paste(min(ds$Engine.rpm), max(ds$Engine.rpm)))
print(paste(min(ds$Lub.oil.pressure), max(ds$Lub.oil.pressure)))
print(paste(min(ds$Fuel.pressure), max(ds$Fuel.pressure)))
print(paste(min(ds$Coolant.pressure), max(ds$Coolant.pressure)))
print(paste(min(ds$lub.oil.temp), max(ds$lub.oil.temp)))
print(paste(min(ds$Coolant.temp), max(ds$Coolant.temp)))
```
# Save the scaled values removed File as a checkpoint
```{r}
write.csv(ds, "C:/Users/nimal/eclipse-workspace/CourseProjects/syCourseProject/DS/dataset/engine_data_scaled.csv", row.names=FALSE)
```
